Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2023-02-22 thirdPaper/draft/typeConeAssociahedra.tex
thank referee
Reference Rup15 et answerPLMS.txt mis à jour

We first recall the results on categorification of cluster algebras that we will use in the proofs.
\begin{theorem}
\label{thm:categorification}
Let~$\cat$ be the cluster category of a valued quiver of Dynkin type~$A,B,C,D,E,F$ or~$G$. Let~$T$ be a basic cluster-tilting object in~$\cat$, and let~$Q$ be the valued Gabriel quiver of~$\End{\cat}(T)$. Then there is a bijection
\[
\CC: \ind(\cat) \xrightarrow{} \{\textrm{cluster variables in the cluster algebra of~$Q$}\},
\]
expressing cluster variables as Laurent polynomials in the variables of the cluster~$\CC(T)$ and which has the follwing properties.
\begin{itemize}
\item[(i)] \emph{(}\cite[Prop.~3.2]{BuanMarshReinekeReitenTodorov}\emph{)} For any~$X,Y \in \ind(\cat)$, $\CC(X)$ and~$\CC(Y)$ are compatible if and only if~$\cat(X, \susp Y) = 0$.
\item[(ii)] \emph{(}\cite[Prop.~4.3]{FuKeller} and \cite[Prop.~2.2]{Palu}, where the proofs also work over an arbitrary base field\emph{)} For any~$X \in \ind(\cat)$, the~$\b{g}$-vector of~$\CC(X)$ is~$\b{g}([X])$, where we identify~$\Z^n$ and~$K_0(\cat ; T)$ via the isomorphism sending~$(a_1, \dots, a_n)$ to~$\sum_{i=1}^n a_i [T_i]$.
\item[(iii)] \emph{(}\cite[Thm.~7.5]{BuanMarshReinekeReitenTodorov}\emph{)} For any~$X,Y \in \ind(\cat)$, $\CC(X)$ and~$\CC(Y)$ are exchangeable if and only if~$\dim_{\field_X}\cat(X, \susp Y) = \dim_{\field_Y}\cat(X, \susp Y) = 1$.
\item[(iv)]\emph{(}\cite{BuanIyamaReitenScott} and \cite[Thm.~5.1]{BuanMarshReiten-mutation}, adapted to the cluster categories of species of Dynkin type\emph{)} Let~$T'$ be any basic cluster-tilting object of~$\cat$. Then the matrix~$B'$ of the associated seed~$(\CC(T'),B')$ is given by the multiplicities of the middle terms of the exchange triangles. 
\end{itemize}
\end{theorem}

@@

We thank H.~Thomas for comments on a preliminary version of the paper and for informing us about the upcoming extended version of~\cite{BazierMatteDouvilleMousavandThomasYildirim}.
We are grateful to C.~Stump for reporting a typo in \cref{exm:typeConeCA}.

>>>>

We first recall the results on categorification of cluster algebras that we will use in the proofs.
\begin{theorem}
\label{thm:categorification}
Let~$\cat$ be the cluster category of a valued quiver of Dynkin type~$A,B,C,D,E,F$ or~$G$. Let~$T$ be a basic cluster-tilting object in~$\cat$, and let~$Q$ be the valued Gabriel quiver of~$\End{\cat}(T)$. Then there is a bijection
\[
\CC: \ind(\cat) \xrightarrow{} \{\textrm{cluster variables in the cluster algebra of~$Q$}\},
\]
expressing cluster variables as Laurent polynomials in the variables of the cluster~$\CC(T)$ and which has the follwing properties.
\begin{itemize}
\item[(i)] \emph{(}\cite[Prop.~3.2]{BuanMarshReinekeReitenTodorov}\emph{)} For any~$X,Y \in \ind(\cat)$, $\CC(X)$ and~$\CC(Y)$ are compatible if and only if~$\cat(X, \susp Y) = 0$.
\item[(ii)] \emph{(}\cite[Prop.~4.3]{FuKeller} and \cite[Prop.~2.2]{Palu}, where the proofs also work over an arbitrary base field\emph{)} For any~$X \in \ind(\cat)$, the~$\b{g}$-vector of~$\CC(X)$ is~$\b{g}([X])$, where we identify~$\Z^n$ and~$K_0(\cat ; T)$ via the isomorphism sending~$(a_1, \dots, a_n)$ to~$\sum_{i=1}^n a_i [T_i]$.
\item[(iii)] \emph{(}\cite[Thm.~7.5]{BuanMarshReinekeReitenTodorov}\emph{)} For any~$X,Y \in \ind(\cat)$, $\CC(X)$ and~$\CC(Y)$ are exchangeable if and only if~$\dim_{\field_X}\cat(X, \susp Y) = \dim_{\field_Y}\cat(X, \susp Y) = 1$.
\item[(iv)]\emph{(}\cite{BuanIyamaReitenScott}, \cite[Thm.~5.1]{BuanMarshReiten-mutation} and \cite[Thm.~5.7]{Rupel15}\emph{)} Let~$T'$ be any basic cluster-tilting object of~$\cat$. Then the matrix~$B'$ of the associated seed~$(\CC(T'),B')$ is given by the multiplicities of the middle terms of the exchange triangles. 
\end{itemize}
\end{theorem}

@@

We thank H.~Thomas for comments on a preliminary version of the paper and for informing us about the upcoming extended version of~\cite{BazierMatteDouvilleMousavandThomasYildirim}.
We are grateful to C.~Stump for reporting a typo in \cref{exm:typeConeCA}.
We also thank an anonymous referee for relevant comments.
